Video-Agnostic Perturbations: Efficient Targeted Attacks \\ for Siamese Visual Tracking

### 2021-01-22 博士/论文提交/adversarial_attack/IJCAI/FormattingGuidelines-IJCAI-21/main.tex
最终版

Given an arbitrary detected or annotated object of interest in the initial video frame, visual object tracking is aimed at {\it recognizing} and {\it localizing} other instances of the same object in subsequent frames. This paradigm of tracking visual objects from a single initial exemplar in the testing phase has been broadly cast as a Siamese network-based one-shot problem recently \cite{SiamFC,SiamRPN,SiamRPN++,SiamFC++}, which is termed as Siamese visual tracking and recognised as highly effective and efficient for visual tracking.

\begin{figure}[htbp]
\centering
%\subfigure{\includegraphics[width=0.2\textwidth]{images/x.jpg}} \qquad
%\subfigure{\includegraphics[width=0.2\textwidth]{images/z.jpg}}
\includegraphics[width=0.48\textwidth]{images/1_v4.pdf}
\caption{Our approach generates the universal perturbations which can be used to attack the Siamese tracker on any video in a targeted manner. Being universal, the generated perturbations can be conveniently exploited to perturb videos on-the-fly without extra computations.}
\label{fig:1}
\end{figure}

@@

Universal adversarial perturbations (UAPs) proposed in \cite{UAP} can fool most images from a data distribution in an image-agnostic manner. Being universal, UAPs can be conveniently exploited to perturb unseen data on-the-fly without extra computations. Therefore, UAPs are particularly useful when attacking real-time applications deployed on platforms with limited computational resources. However, no existing work has touched the topic of attacking the Siamese trackers using UAPs, because it is hard to apply existing UAPs to attack Siamese trackers directly. The main reason lies in the fact that, (a) most UAPs are designed for typical neural networks with one image as input while Siamese networks accept both the template and search images, and (b) the goal of existing UAP methods is to disturb unary and binary model outputs for single instance while we need to use universal perturbations to mislead Siamese trackers to follow a specified trajectory.

In this paper, we make the first attempt in finding the video-agnostic perturbations that fool state-of-the-art Siamese trackers in a targeted attack manner, which comes at virtually no cost in the online-tracking phase. Specifically, we aim to attack the trackers by adding a universal imperceptible perturbation to the template image and pasting a \textit{fake target}, \ie, a small universal adversarial patch, into the search image adhering to the predefined trajectory, so that the tracker outputs the location and size of the \textit{fake target} instead of the real target. Our generated video-agnostic perturbations allow perturbing a novel video to come at no additional cost except the mere addition and pasting operations -- and not require gradient optimization or network inference. Experiment results on OTB2015 \cite{OTB}, GOT-10k \cite{GOT-10k} and LaSOT \cite{GOT-10k} datasets demonstrate the effectiveness and efficiency of our approach (see Figure \ref{fig:1}).

@@

Siamese visual tracking is a fundamental research direction in template matching-based tracking besides the correlation filter-based methods. Both of them are aimed to estimate the positions of a template cropped from the initial video frame ``causally" in the subsequent frames. Siamese trackers formulate visual tracking as learning cross-correlation similarities between a target template and the candidates in search region in a convolution fashion. Tracking is then performed by locating the object in the search image region based on the highest visual similarity. This paradigm is formulated as a local one-shot detection task.

@@

\textit{Imperceptible Perturbations v.s. Adversarial Patch.} The imperceptible perturbations most commonly modify each pixel by a small amount and can be found using a number of optimisation strategies such as Limited-memory BFGS \cite{intriguing} and PGD \cite{PGD}.
Different from the imperceptible perturbations, the adversarial patch is extremely salient to a neural network. The adversarial patch can be placed anywhere into the input image to cause the network to misbehave, and thus is commonly used for universal attacks \cite{patch}.
To the best of our knowledge, we are the first to attack object trackers utilizing both the imperceptible perturbation and the adversarial patch together, which are jointly trained in an end-to-end manner.
Note that our adversarial patch works in the network domain instead of the image domain. In the network-domain case, the noise is allowed to take any value and is not restricted to the dynamic range of images as in the image-domain case \cite{karmon2018lavan}.

\textit{Non-targeted Attacks v.s. Targeted Attacks.} In the case of non-targeted attacks, the adversary's goal is to cause the network to predict any incorrect label and the specific incorrect label does not matter, \eg, pushing the object position estimation just outside the true search region in visual tracking.
Targeted attacks, however, aim to change the network's prediction to some specific target label. In visual tracking, the targeted attacks aim to intentionally drive trackers to output specified object positions following a pre-defined trajectory.

@@

Recently, there are several explorations of the adversarial attacks to the visual tracking task. For example, PAT \cite{PAT} generates physical adversarial textures via white-box attacks to steer the tracker to lock on the texture when a tracked object moves in front of it. However, PAT validates its method by attacking a light deep regression tracker GOTURN \cite{GOTURN}, which has low tracking accuracy on modern benchmarks. In this paper, we aim to attack the state-of-the-art Siamese trackers.
RTAA \cite{RTAA} takes temporal motion into consideration when generating lightweight perturbations over the estimated tracking results frame-by-frame. However, RTAA only performs the non-targeted attacks for trackers, which is less challenging than the targeted attacks in this paper, as we aim to create arbitrary, complex trajectories at test time.

Targeted attacks to follow an erroneous path which looks realistic are crucial to deceive the tracking system without raising any suspicion in the real-world applications.
SPARK \cite{SPARK} computes incremental perturbations by using information from the past frames to perform targeted attacks to Siamese trackers. However, SPARK needs to generate distinct adversarial example for every search image through heavy iterative schemes, which is time-consuming to attack online-tracking in real time. The recent real-time attacker in \cite{TTP} exclusively uses the template image to generate temporally-transferable perturbations in a one-shot manner, and then adds them to every search image. However, this method still needs to generate perturbations for each individual video, and its targeted attack setting requires diverse perturbations from several runs of network inference. It is ill-suited to attack a real-world online-tracking system when we can not get access to the limited computational resources. In this paper, however, we propose video-agnostic perturbations which allow perturbing a novel video to come at no additional cost except the mere addition and pasting operations.

\begin{figure}[t]
\centering
\includegraphics[width=0.48\textwidth]{images/network_v4.png}
\caption{The pipeline of the proposed method. We aim to train an imperceptible perturbation $\delta$ for the template image $\textbf z$, and an adversarial patch $p$ for the search image $\textbf x$. After adding $\delta$ to $\textbf z$ and pasting the fake target $p$ to $\textbf x$, the tracker outputs the location and size of the adversarial patch instead of the real target.}
\label{fig:net}
\end{figure}

@@

\textcolor{black}{In this section, we introduce our universal targeted attack framework for object trackers based on Siamese networks. We aim to attack the tracker by adding an imperceptible perturbation to the template image and pasting a \textit{fake target}, i.e., an adversarial patch into the search image adhering to the predefined trajectory, so that the tracker outputs the location and size of the fake target instead of the real target. Below, we formalize the task of performing universal targeted attacks on an object tracker, and then introduce our perturbation strategy in detail.}

@@

Let $V=\{I_i\}_1^T$ denote the frames of a video sequence of length $T$.
$B^{gt}=\{b^{gt}_i\}_1^T$ is used to represent the target's ground-truth position in each frame.
The visual object tracking will predict the position $B^{pred}=\{b^{pred}_i\}_1^T$ of this target in the subsequent frames when given its initial state.
In SiamFC++ \cite{SiamFC++}, the tracker first transforms the reference frame $I_1$ and annotation $b_1^{gt}$ to get an template image $\textbf z$, and transforms the candidate frame $I_i$ to get the search image $\textbf x_i$ centered at the position obtained in the previous frame.
At each time-step, the template image $\textbf z$ and the search image $\textbf x_i$ are first passed individually through a shared backbone network, and the resulting features are processed by some non-shared layers and fused by depth-wise separable correlation. The fused features then act as input to a head network, which predicts a classification map $\textbf{C}$, a bounding box regression map $\textbf{R}$, and a quality assessment map $\textbf{Q}$. In short, $\textbf C$ encodes the probability of each spatial position to contain the target, $\textbf R$ regresses the bounding box position of the target, and $\textbf Q$ predicts the target state estimation quality. The final bounding box is then generated according to $\textbf{C}$, $\textbf{R}$ and $\textbf{Q}$.

Formally, we aim to train an imperceptible perturbation $\delta$ for the template image $\textbf z$, and an adversarial patch $p$ for the search image $\textbf x$. After adding $\delta$ to $\textbf z$ and pasting the fake target $p$ to $\textbf x$, the tracker outputs the location and size of the adversarial patch instead of the real target (see Figure \ref{fig:net}).
Both $\delta$ and $p$ are universal (i.e., video-agnostic), which means perturbing a new video only involves the mere addition/paste of the perturbations to the template/search image -- and does not require gradient optimisation or network inference.

\begin{algorithm}[tb]
\caption{Training Process}
\label{alg:algorithm}
\textbf{Input}: Training dataset $\mathcal{V}$, Siamese tracker $\phi$, and max iteration number $N$.\\
\textbf{Output}: Universal perturbation $\delta$, and universal patch $p$.
\begin{algorithmic}[1] %[1] enables line numbers
\STATE Let $k = 0$.
\WHILE{$k < N$}
\STATE Randomly pick a video $V\in \mathcal{V}$. The corresponding ground truth is $B^{gt}=\{b^{gt}_i\}^T_1$.
\STATE Randomly pick two frames $I_t, I_s$ from $V$.
\STATE Generate template image $\textbf{z}$ according to $I_t$ and $b^{gt}_t$.
\STATE $\tilde{\textbf{z}} = \textbf{z} + \delta_k.$
\STATE Generate search image $\textbf{x}$ according to $I_s$ and $b^{gt}_s$.
\STATE Calculate the fake target position $\{l^x, l^y, w, h\}$ with respect to the search image.
\STATE $\tilde{\textbf x} = A(\textbf x, p_k, (l^x, l^y), (w, h)).$
\STATE $\textbf{C, R, Q} = \phi(\tilde {\textbf x}, \tilde{\textbf z}).$
\STATE Generate fake labels $\textbf{C}^*,\textbf{R}^*,\textbf{Q}^*$ using $\{l^x, l^y, w, h\}$.
\STATE Calculate loss $L(\textbf{C, R, Q}, \textbf{C}^*, \textbf{R}^*, \textbf{Q}^*)$ using Equ. \ref{eq:loss}.
\STATE $\delta_{k+1} = \delta_{k} - \epsilon_1 \cdot \text{sign}(\nabla_{\delta_k}L).$
\STATE $p_{k+1} = p_{k} - \epsilon_2 \cdot \text{sign}(\nabla_{p_k}L).$
\STATE $k = k + 1.$
\ENDWHILE
\STATE \textbf{return} $\delta_N, p_N.$
\end{algorithmic}
\label{alg}
\end{algorithm}

\begin{figure*}[t]
\centering
\includegraphics[width=0.95\textwidth]{images/vis_v4.pdf}
\caption{The results under targeted attacks. Red represents a ground-truth bounding box, yellow represents the targeted bounding box, and green represents the predicted bounding box by the tracker. The yellow and green boxes are basically the same in time series, which indicates that the targeted attack is successful.}
\label{fig:vis}
\end{figure*}

\subsection{Generating Universal Perturbations}

In this subsection, we show how to train the universal perturbations $(\delta, p)$ for Siamese trackers (see Algorithm \ref{alg}).
During the $k$-th iteration of training, a video $V=\{I_i\}_1^T$ is randomly selected from the training dataset $\mathcal V$. Assuming the template perturbation at the $k$-th iteration is $\delta_k \in \mathbb{R}^{127\times 127 \times 3}$, and the adversarial patch is $p_k \in \mathbb{R}^{128\times 128\times 3}$. We first randomly pick two frames $I_t, I_s$ from $V$.
The clean template image $\textbf z\in \mathbb{R}^{127\times 127 \times 3}$ is generated according to $I_t$ and $b^{gt}_t$, and the perturbed template image is:
\begin{equation}
\tilde {\textbf z} = \textbf z + \delta_i.
\end{equation}
Similarly, the clean search image $\textbf x \in \mathbb{R}^{303\times 303 \times 3}$ is generated according to $I_s$ and $b^{gt}_s$.
As mentioned before, the patch is regarded as a fake target and pasted on the search image. We force the center position of the fake target to near the center position of the real target with a shift range of 64 pixels, where shift is defined as the max range of translation generated by a uniform distribution.
The width/height of the fake target is randomly selected between 32 pixels and 128 pixels.
The perturbed search image is generated as follows:
\begin{equation}
\tilde{\textbf x} = A(\textbf x, p_k, (l^x, l^y), (w, h)),
\end{equation}
where $(l^x, l^y)$ and $(w, h)$ represents the position and size of the fake target with respect to the search image, respectively. $A$ is \textit{patch application operator} \cite{patch} which first resizes the patch $p_k \in \mathbb{R}^{128\times 128\times 3}$ to $\hat{p}_k \in \mathbb{R}^{w\times h\times 3}$, and then applies the resized patch $\hat{p}_k$ to the search image $\textbf x$ at location $(l^x,l^y)$.

Subsequently, the SiamFC++ tracker $\phi(\cdot)$ takes $\tilde {\textbf x}$ and $\tilde{\textbf  z}$ as input and makes predictions as follows:
\begin{equation}
\textbf{C, R, Q} = \phi(\tilde {\textbf x}, \tilde{\textbf z}),
\end{equation}

\subsubsection{Training Objective}

The loss function is calculated as follows:
\begin{equation}
\begin{array}{l}
\begin{aligned}
L&=\frac{\alpha}{N_{\mathrm{pos}}} \sum_{x, y} L_{\mathrm{cls}}\left(\textbf{C}_{x, y}, \textbf{C}_{x, y}^{*}\right) \\
&+\frac{\beta}{N_{\mathrm{pos}}} \sum_{x, y} \textbf{1}_{\left\{\textbf{C}_{x, y}^{*}>0\right\}} L_{\mathrm{quality}}\left(\textbf{Q}_{x, y}, \textbf{Q}_{x, y}^{*}\right) \\
&+\frac{\gamma}{N_{\mathrm{pos}}} \sum_{x, y} \textbf{1}_{\left\{\textbf{C}_{x, y}^{*}>0\right\}} L_{\mathrm{reg}}\left(\textbf{R}_{x, y}, \textbf{R}_{x, y}^{*}\right) \\
&+\eta \cdot ||\delta_k||_2^2 +  \sigma \cdot ||p_k||^2_2,
\end{aligned}
\end{array}
\label{eq:loss}
\end{equation}
\textcolor{black} %(SiamFC++)
{where $\textbf{C}_{x, y}, \textbf{R}_{x, y}, \textbf{Q}_{x, y}$ represent the values of $\textbf{C}, \textbf{R}, \textbf{Q}$ at location $(x, y)$, respectively. $\textbf{C}^*, \textbf{R}^*, \textbf{Q}^*$ are the fake labels generated according to the position/size of the fake target.$\textbf 1$ is the indicator function that takes 1 if the condition in subscribe holds and takes 0 if not, $L_{\mathrm{cls}}$ denotes the focal loss \cite{focal} for classification result, $L_{\mathrm{quality}}$ denotes the binary cross entropy (BCE) loss for quality assessment and $L_{\mathrm{reg}}$ denotes the IoU loss \cite{iou-loss} for bounding box result. Following SiamFC++, we assign 1 to $\textbf{C}_{x, y}^{*}$ if $(x, y)$ is considered as a positive sample, and 0 if as a negative sample.}

\subsubsection{Optimisation}

At each training step, the perturbations are updated as follows:
\begin{gather}
\delta_{k+1} = \delta_{k} - \epsilon_1 \cdot \text{sign}(\nabla_{\delta_k}L)\\
p_{k+1} = p_{k} - \epsilon_2 \cdot \text{sign}(\nabla_{p_k}L),
\end{gather}
where $\epsilon_1$ is used to ensure that the perturbation added to the template image is imperceptible, and $\epsilon_2$ is used to maintain the training stability.
During training, we only optimize the values of perturbations $(\delta, p)$, and the parameters of the Siamese network remain intact.

@@

Once the perturbations $(\delta, p)$ is trained, we can use them to perturb the template image and search images of any natural video. Both $\delta$ and $p$ are universal (i.e., video-agnostic), which means perturbing a new video only involves the mere addition/paste of the perturbations to the template/search image -- and does not require gradient optimisation or network inference.
Assume $B^{fake}=\{b^{fake}_i\}_1^{T}$ is the trajectory we hope the tracker to output.
During tracking the $i$-th frame of the video $V=\{I_i\}_1^T$, we need to transform the bounding box $b^{fake}_i$ (with respect to the original frame $I_i$) into the box $\hat b_i=\{l^x_i, l^y_i, w_i, h_i\}$ with respect to the search image $\textbf x_i$, and paste $p$ into $\textbf x_i$ according to $\hat b_i$:
\begin{equation}
\tilde{\textbf x}_i = A(\textbf x_i, p, (l^x_i, l^y_i), (w_i, h_i)).
\end{equation}
The tracker then takes $\tilde{\textbf z}_i=\textbf z_i+\delta$ and $\tilde{ \textbf x}_i$ as input, and the subsequent tracking procedure remains the same as SiamFC++.

@@

\textit{Transferability to Different Backbones} We evaluate the transferability of our attacks when applying the perturbations to different backbones of SiamFC++, \ie, ShuffleNet \cite{ShuffleNet} and AlexNet \cite{AlexNet}.
The experimental results are shown in Table \ref{tab:backbone}. In the case of SiamFC++-AlexNet, the AO with respect to the \textit{real trajectory} drops from 0.72 to 0.196. However, our perturbations do not generalize well to SiamFC++-ShuffleNet, which we conjecture to be due to the particular group convolution and channel shuffle operations in ShuffleNet.

>>>>

Given an arbitrary detected or annotated object of interest in the initial video frame, visual object tracking is aimed at {\it recognizing} and {\it localizing} other instances of the same object in subsequent frames. This paradigm of tracking visual objects from a single initial exemplar in the online-tracking phase has been broadly cast as a Siamese network-based one-shot problem recently \cite{SiamFC,SiamRPN,SiamRPN++,SiamFC++}, which is termed as Siamese visual tracking and recognised as highly effective and efficient for visual tracking.

\begin{figure}[htbp]
\centering
%\subfigure{\includegraphics[width=0.2\textwidth]{images/x.jpg}} \qquad
%\subfigure{\includegraphics[width=0.2\textwidth]{images/z.jpg}}
\includegraphics[width=0.45\textwidth]{images/1_v8.pdf}
\caption{An illustration of our attacks to SiamFC++ on some example tracking sequences from GOT-10k benchmark. Our approach generates the video-agnostic perturbations which can force SiamFC++ to follow a complicated trajectory at virtually no cost. This is realized by off-line training the perturbations so that the tracker mistakenly believes the \textit{fake target} area contains the object to be tracked (see the left heatmap). Moreover, the quality assessment branch in SiamFC++ is also misled to confirm this result (see the right heatmap). The \textit{fake target} size is gradually decreased.} 
%Being universal, the generated perturbations can be conveniently exploited to perturb videos on-the-fly without extra computations. 
%Red represents the ground-truth bounding box, and green represents the predicted bounding box by the tracker.
%The purple dashed box marks the universal adversarial patch. The left heatmap represents the probability of each spatial position to contain the \textit{fake target}, and the right heatmap represents the target state estimation quality.}
\label{fig:1}
\end{figure}

@@

Universal adversarial perturbations (UAPs) proposed in \cite{UAP} can fool most images from a data distribution in an image-agnostic manner. Being universal, UAPs can be conveniently exploited to perturb unseen data on-the-fly without extra computations. Therefore, UAPs are particularly useful when attacking real-time applications deployed on platforms with limited computational resources. However, no existing work has touched the topic of attacking the Siamese trackers using UAPs, because it is hard to apply existing UAPs to attack Siamese trackers directly. The main reason lies in the fact that, (a) most UAPs are designed for typical neural networks with one image as input while Siamese networks accept both the template and search images, and (b) the goal of existing UAP methods is to disturb unary or binary model outputs for single instance while we need to use universal perturbations to mislead Siamese trackers to follow a specified trajectory.

In this paper, we make the first attempt in finding the video-agnostic perturbations that fool a state-of-the-art Siamese tracker, \ie, SiamFC++ \cite{SiamFC++}, in a targeted attack manner, which comes at virtually no cost in the online-tracking phase. Specifically, we aim to attack the trackers by adding a universal imperceptible perturbation to the template image and pasting a \textit{fake target}, \ie, a small universal adversarial patch, into the search image adhering to the predefined trajectory (as shown in Figure \ref{fig:1}), so that the tracker outputs the location and size of the \textit{fake target} instead of the real target. Our generated video-agnostic perturbations allow perturbing a novel video to come at no additional cost except the mere addition and pasting operations -- and not require gradient optimization or network inference. Experiment results on OTB2015 \cite{OTB}, GOT-10k \cite{GOT-10k} and LaSOT \cite{GOT-10k} benchmarks demonstrate the effectiveness and efficiency of our approach.

@@

Siamese visual tracking is a fundamental research direction in template matching-based tracking besides the correlation filter-based methods. Both of them are aimed to "causally" estimate the positions of a template cropped from the initial video frame in the subsequent frames. Siamese trackers formulate visual tracking as learning cross-correlation similarities between a target template and the candidates in search region in a convolution fashion. Tracking is then performed by locating the object in the search image region based on the highest visual similarity. This paradigm is formulated as a local one-shot detection task.

@@

\textit{Imperceptible Perturbations v.s. Adversarial Patch} The imperceptible perturbations most commonly modify each pixel by a small amount and can be found using a number of optimisation strategies such as Limited-memory BFGS \cite{intriguing} and PGD \cite{PGD}.
Different from the imperceptible perturbations, the adversarial patch is extremely salient to a neural network. The adversarial patch can be placed anywhere into the input image to cause the network to misbehave, and thus is commonly used for universal attacks \cite{patch}.
To the best of our knowledge, we are the first to attack object trackers utilizing both the imperceptible perturbation and the adversarial patch together, which are jointly trained in an end-to-end manner.
Note that our adversarial patch works in the network domain instead of the image domain. In the network-domain case, the noise is allowed to take any value and is not restricted to the dynamic range of image value as in the image-domain case \cite{karmon2018lavan}.

\textit{Untargeted Attacks v.s. Targeted Attacks} In the case of untargeted attacks, the adversary's goal is to cause the network to predict any incorrect label and whatever the incorrect label is does not matter, \eg, pushing the object location estimation just outside the true search region in visual tracking.
Targeted attacks, however, aim to change the network's prediction to some specific target label. In visual tracking, the targeted attacks aim to intentionally drive trackers to output specified object locations following a predefined trajectory.

@@

Recently, there are several explorations of the adversarial attacks to the visual tracking task. For example, PAT \cite{PAT} generates physical adversarial textures via white-box attacks to steer the tracker to lock on the texture when a tracked object moves in front of it. However, PAT validates its method by attacking a light deep regression tracker GOTURN \cite{GOTURN}, which has low tracking accuracy on modern benchmarks. In this paper, we aim to attack the state-of-the-art Siamese trackers.
RTAA \cite{RTAA} takes temporal motion into consideration when generating lightweight perturbations over the estimated tracking results frame-by-frame. However, RTAA only performs the untargeted attacks for trackers, which is less challenging than the targeted attacks in this paper, as we aim to create arbitrary, complex trajectories at test time.

Targeted attacks to follow an erroneous path which looks realistic are crucial to deceive the tracking system without raising any suspicion in the real-world applications.
SPARK \cite{SPARK} computes incremental perturbations by using information from the past frames to perform targeted attacks to Siamese trackers. However, SPARK needs to generate distinct adversarial example for every search image through heavy iterative schemes, which is time-consuming to attack online-tracking in real time. The recent real-time attacker in \cite{TTP} exclusively uses the template image to generate temporally-transferable perturbation in a one-shot manner, and then adds it to every search image. However, this method still needs to generate perturbations for each individual video, and its targeted attack setting requires diverse perturbations from several runs of network inference. It is ill-suited to attack a real-world online-tracking system when we can not get access to the limited computational resources. In this paper, however, we propose video-agnostic perturbations which allow perturbing a novel video to come at no additional cost except the mere addition and pasting operations.

\begin{figure}[t]
\centering
\includegraphics[width=0.48\textwidth]{images/network_v5.pdf}
\caption{The training pipeline of the proposed method. We aim to train an imperceptible perturbation $\delta$ for the template image $\textbf z$, and an adversarial patch $p$ for the search image $\textbf x$. After adding $\delta$ to $\textbf z$ and pasting the \textit{fake target} $p$ into $\textbf x$, the tracker outputs the location and size of the \textit{fake target} instead of the real target.}
\label{fig:net}
\end{figure}

@@

\textcolor{black}{In this section, we introduce our video-agnostic targeted attack framework for Siamese trackers. We aim to attack the tracker by adding an imperceptible perturbation to the template image and pasting a \textit{fake target}, i.e., an adversarial patch, into the search images adhering to the predefined trajectory, so that the tracker outputs the location and size of the \textit{fake target} instead of the real target. Below, we formalize our targeted attacks to SiamFC++ \cite{SiamFC++}, and then introduce our perturbation strategy in detail.}

@@

Let $V=\{I_i\}_1^T$ denote the frames of a video sequence of length $T$.
$B^{gt}=\{b^{gt}_i\}_1^T$ is used to represent the target's ground-truth position in each frame.
The visual object tracking aims to predict the position $B^{pred}=\{b^{pred}_i\}_1^T$ of this target in the subsequent frames given its initial state.
In SiamFC++, the tracker first transforms the paired reference frame $I_1$ and annotation $b_1^{gt}$ to get an template image $\textbf z$, and transforms the search frame $I_i$ to get the search image $\textbf x_i$ centered at the position estimated in the previous frame.
At each time-step, the template image $\textbf z$ and the search image $\textbf x_i$ are first passed individually through a shared backbone network, and the resulting features are processed by some non-shared layers and fused by depth-wise separable correlation. The fused features then act as input to a head network, which predicts a classification map $\textbf{C}$, a bounding box regression map $\textbf{R}$, and a quality assessment map $\textbf{Q}$ in an anchor-free manner. In short, $\textbf C$ encodes the probability of each spatial position to contain the target, $\textbf R$ regresses the bounding box of the target, and $\textbf Q$ predicts the target state estimation quality. The final bounding box is then generated according to $\textbf{C}$, $\textbf{R}$ and $\textbf{Q}$.

Formally, we aim to train an imperceptible perturbation $\delta$ for the template image $\textbf z$, and an adversarial patch $p$ for the search image $\textbf x_i$. After adding $\delta$ to $\textbf z$ and pasting the fake target $p$ into $\textbf x_i$, the tracker outputs the location and size of the adversarial patch instead of the real target (see Figure \ref{fig:net}).
Both $\delta$ and $p$ are universal (\ie, video-agnostic), which means perturbing a novel video only involves the mere addition and pasting of the perturbations to the template and search images -- and does not require gradient optimisation or network inference.

\subsection{Generating Video-Agnostic Perturbations}

In this subsection, we show how to train the video-agnostic perturbations $(\delta, p)$ for Siamese trackers.
During the $k$-th iteration of training, a video $V=\{I_i\}_1^T$ is randomly selected from the training dataset $\mathcal V$. Assuming the template perturbation at the $k$-th iteration is $\delta_k \in \mathbb{R}^{127\times 127 \times 3}$, and the adversarial patch is $p_k \in \mathbb{R}^{128\times 128\times 3}$. We first randomly pick paired frames $I_t, I_s$ from $V$.
The clean template image $\textbf z\in \mathbb{R}^{127\times 127 \times 3}$ is generated according to $I_t$ and $b^{gt}_t$, and the perturbed template image is:
\begin{equation}
\tilde {\textbf z} = \textbf z + \delta_k.
\end{equation}
Similarly, the clean search image $\textbf x \in \mathbb{R}^{303\times 303 \times 3}$ is generated according to $I_s$ and $b^{gt}_s$.
As mentioned before, the patch is regarded as a \textit{fake target} and pasted into the search images. We force the center position of the \textit{fake target} to near the center position of the real target within a shift range of 64 pixels, where shift is defined as the maximum range of translation generated from a uniform distribution.
The width/height of the \textit{fake target} is randomly selected between 32 pixels and 128 pixels.
The perturbed search image is generated as follows:
\begin{equation}
\tilde{\textbf x} = A(\textbf x, p_k, (l^x, l^y), (w, h)),
\end{equation}
where $(l^x, l^y)$ and $(w, h)$ represent the position and size of the \textit{fake target} with respect to the search image, respectively. $A$ is a patch application operator \cite{patch} which first resizes the patch $p_k \in \mathbb{R}^{128\times 128\times 3}$ to $\hat{p}_k \in \mathbb{R}^{w\times h\times 3}$, and then pastes the resized patch $\hat{p}_k$ into the search image $\textbf x$ at location $(l^x,l^y)$.

Subsequently, the SiamFC++ tracker $\phi(\cdot)$ takes $\tilde {\textbf x}$ and $\tilde{\textbf  z}$ as input and makes predictions as follows:
\begin{equation}
\textbf{C, R, Q} = \phi(\tilde {\textbf x}, \tilde{\textbf z}).
\end{equation}

\textit{Training Objective} The loss function is calculated as follows:
\begin{equation}
\begin{array}{l}
\begin{aligned}
L&=\frac{\alpha}{N_{\mathrm{pos}}} \sum_{x, y} L_{\mathrm{cls}}\left(\textbf{C}_{x, y}, \textbf{C}_{x, y}^{*}\right) \\
&+\frac{\beta}{N_{\mathrm{pos}}} \sum_{x, y} \textbf{1}_{\left\{\textbf{C}_{x, y}^{*}>0\right\}} L_{\mathrm{quality}}\left(\textbf{Q}_{x, y}, \textbf{Q}_{x, y}^{*}\right) \\
&+\frac{\gamma}{N_{\mathrm{pos}}} \sum_{x, y} \textbf{1}_{\left\{\textbf{C}_{x, y}^{*}>0\right\}} L_{\mathrm{reg}}\left(\textbf{R}_{x, y}, \textbf{R}_{x, y}^{*}\right) \\
&+\eta \cdot ||\delta_k||_2^2 +  \sigma \cdot ||p_k||^2_2,
\end{aligned}
\end{array}
\label{eq:loss}
\end{equation}
\textcolor{black} %(SiamFC++)
{where $\textbf{C}_{x, y}, \textbf{R}_{x, y}, \textbf{Q}_{x, y}$ represent the values of $\textbf{C}, \textbf{R}, \textbf{Q}$ at location $(x, y)$, respectively. $\textbf{C}^*, \textbf{R}^*, \textbf{Q}^*$ are the fake labels generated according to the position and size of the \textit{fake target}. $\textbf 1$ is the indicator function that takes 1 if the condition in subscribe holds and takes 0 if not, $N_{\mathrm{pos}}$ denotes the number of positive samples in the training phase, $L_{\mathrm{cls}}$ denotes the focal loss \cite{focal} for classification result, $L_{\mathrm{quality}}$ denotes the binary cross entropy (BCE) loss for quality assessment, and $L_{\mathrm{reg}}$ denotes the IoU loss \cite{iou-loss} for bounding box regression. Following SiamFC++, we assign 1 to $\textbf{C}_{x, y}^{*}$ if $(x, y)$ is considered as a positive sample, and 0 if as a negative sample.}

\begin{algorithm}[tb]
\caption{Training Process}
\label{alg:algorithm}
\textbf{Input}: Training dataset $\mathcal{V}$, Siamese tracker $\phi$, and max iteration number $N$.\\
\textbf{Output}: Imperceptible perturbation $\delta$, and adversarial patch $p$.
\begin{algorithmic}[1] %[1] enables line numbers
\STATE Let $k = 0$.
\WHILE{$k < N$}
\STATE Randomly pick a video $V\in \mathcal{V}$. The corresponding ground truth is $B^{gt}=\{b^{gt}_i\}^T_1$.
\STATE Randomly pick paired frames $I_t, I_s$ from $V$.
\STATE Generate template image $\textbf{z}$ according to $I_t$ and $b^{gt}_t$.
\STATE $\tilde{\textbf{z}} = \textbf{z} + \delta_k.$
\STATE Generate search image $\textbf{x}$ according to $I_s$ and $b^{gt}_s$.
\STATE Calculate the \textit{fake target} position $\{l^x, l^y, w, h\}$ with respect to the search image.
\STATE $\tilde{\textbf x} = A(\textbf x, p_k, (l^x, l^y), (w, h)).$
\STATE $\textbf{C, R, Q} = \phi(\tilde {\textbf x}, \tilde{\textbf z}).$
\STATE Generate fake labels $\textbf{C}^*,\textbf{R}^*,\textbf{Q}^*$ using $\{l^x, l^y, w, h\}$.
\STATE Calculate loss $L(\textbf{C, R, Q}, \textbf{C}^*, \textbf{R}^*, \textbf{Q}^*)$ using Equ. \ref{eq:loss}.
\STATE $\delta_{k+1} = \delta_{k} - \epsilon_1 \cdot \text{sign}(\nabla_{\delta_k}L).$
\STATE $p_{k+1} = p_{k} - \epsilon_2 \cdot \text{sign}(\nabla_{p_k}L).$
\STATE $k = k + 1.$
\ENDWHILE
\STATE \textbf{return} $\delta_N, p_N.$
\end{algorithmic}
\label{alg}
\end{algorithm}

\begin{figure*}[t]
\centering
\includegraphics[width=0.95\textwidth]{images/vis_v4.pdf}
\caption{Our targeted attack results following a predefined \textit{fake trajectory} (indicated by the yellow targeted bounding box).}
\label{fig:vis}
\end{figure*}

\textit{Optimisation} At each training step, the perturbations are updated as follows:
\begin{gather}
\delta_{k+1} = \delta_{k} - \epsilon_1 \cdot \text{sign}(\nabla_{\delta_k}L)\\
p_{k+1} = p_{k} - \epsilon_2 \cdot \text{sign}(\nabla_{p_k}L),
\end{gather}
where $\epsilon_1$ is used to ensure that the perturbation added to the template image is imperceptible, and $\epsilon_2$ is used to maintain the training stability.
During training, we only optimize the values of perturbations $(\delta, p)$ and the parameters of the Siamese network remain intact.

We outline this training procedure in Algorithm \ref{alg}.

@@

Once the perturbations $(\delta, p)$ is trained, we can use them to perturb the template image and search images of any novel video for attacking. Both $\delta$ and $p$ are universal (i.e., video-agnostic), which means perturbing a novel video only involves the mere addition and pasting of the perturbations to the template and search images -- and does not require gradient optimisation or network inference.
Assume $B^{fake}=\{b^{fake}_i\}_1^{T}$ is the trajectory we hope the tracker to output.
During tracking the $i$-th frame of the video $V=\{I_i\}_1^T$, we need to transform the bounding box $b^{fake}_i$ (with respect to the original frame $I_i$) into the box $\hat b_i=\{l^x_i, l^y_i, w_i, h_i\}$ with respect to the search image $\textbf x_i$, and paste $p$ into $\textbf x_i$ according to $\hat b_i$:
\begin{equation}
\tilde{\textbf x}_i = A(\textbf x_i, p, (l^x_i, l^y_i), (w_i, h_i)).
\end{equation}
The tracker then takes $\tilde{\textbf z}_i=\textbf z_i+\delta$ and $\tilde{ \textbf x}_i$ as input, and the subsequent tracking procedure remains the same as SiamFC++.

@@

\textit{Transferability to Different Backbones} We evaluate the transferability of our attacks when applying the perturbations to two more different backbones of SiamFC++, \ie, ShuffleNet \cite{ShuffleNet} and AlexNet \cite{AlexNet}.
The experimental results are shown in Table \ref{tab:backbone}. In the case of SiamFC++-AlexNet, the AO with respect to the \textit{real trajectory} drops from 0.72 to 0.196. However, our perturbations do not generalize well to SiamFC++-ShuffleNet, which we conjecture to be due to the particular group convolution and channel shuffle operations in ShuffleNet.
